Comparative analysis of PMV Models accuracy implemented in the ISO 7730:2005 and ASHRAE 55:2023

### 2025-01-28 Manuscript/src/sections/conclusions.tex
review: removed ed's comments

We also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} standards.
Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
\todo{EA. But this statement is incorrect for ASHRAE Std 55, which only defines its single comfort zone using +/-0.5}
\todo[inline, color=blue!40]{FT. We clarified this issue above.}
Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
\todo{unclear what limits you are talking about, or who the audience is for your suggestions.  The ComfortTool allows extrapolation to high PMVs, but that is not true for ASHRAE 55 itself.  
Your suggestions are just aimed at you as keepers of the ComfortTool}
\todo[inline, color=blue!40]{FT. The limits are the one described in Table 2, for example, we suggest to use the model only from 0.3 and 1.5 clo and met between 1 and 2.}
The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
Consequently, we suggest reducing both standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
\begin{table}[htb!]
    \centering
    \begin{tabular}{cc}
        \toprule
        Variable & New Proposed range \\
        \midrule
        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
        \ac{rh} & \qtyrange{10}{80}{\percent} \\
        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
        \ac{met} & \qtyrange{1}{2}{met} \\
        \ac{pmv} & \numrange{-.5}{.5} \\
        \bottomrule
    \end{tabular}
    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} standards.}
    \label{tab:ranges}
\end{table}
\mycite{Humphreys2002} reached a similar conclusion and observed that the ranges of inputs consistent with a valid use of PMV are much narrower than those given in \gls{7730}.
The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
We then recommend that both the \gls{55} and \gls{7730} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.

The \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
Assuming that people who are `slightly warm' or `slightly cool' are thermally neutral leads to undercounting thermal discomfort in our dataset, which is not a valid assumption.
We recommend the use of the thermal preference scale to determine how people perceive their thermal environment. 
This scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.
\todo{EA. This is a major suggestion with very significant implications for energy etc that have not been discussed in the paper.  There are surely reasons people have not used that scale on its own, for this purpose.}
\todo[inline, color=blue!40]{FT. We discussed this in Figure 5a and in its related text. The PMV model assume that people are comfortable when they claim to be slightly warm or slightly cool, this is explicit in Fanger's book.}

>>>>

We also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} standards.
Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
Consequently, we suggest reducing both standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
\begin{table}[htb!]
    \centering
    \begin{tabular}{cc}
        \toprule
        Variable & New Proposed range \\
        \midrule
        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
        \ac{rh} & \qtyrange{10}{80}{\percent} \\
        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
        \ac{met} & \qtyrange{1}{2}{met} \\
        \ac{pmv} & \numrange{-.5}{.5} \\
        \bottomrule
    \end{tabular}
    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} standards.}
    \label{tab:ranges}
\end{table}
\mycite{Humphreys2002} reached a similar conclusion and observed that the ranges of inputs consistent with a valid use of PMV are much narrower than those given in \gls{7730}.
The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
We then recommend that both the \gls{55} and \gls{7730} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.

The \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
Assuming that people who are `slightly warm' or `slightly cool' are thermally neutral leads to undercounting thermal discomfort in our dataset, which is not a valid assumption.
We recommend the use of the thermal preference scale to determine how people perceive their thermal environment. 
This scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.

### 2025-01-28 Manuscript/src/sections/introduction.tex
review: removed ed's comments

\todo[inline]{Ed's comment:
The PMV model consists of a heat balance model and an empirical equation that predicts thermal sensation (PMV) from heat balance.  (need to describe this further, especially the simplified way that sweating is treated by the heat balance model).
In 2010 ASHRAE Std 55 modified the PMV heat balance calculation to make it more capable of predicting the cooling caused by elevated air movement under warm conditions.
It uses the Gagge two-node model to compute the convective and evaporative heat loss from the body. This model explicitly models physiological cooling in warm conditions where evaporation of sweat becomes a significant source of heat loss. The model output can be expressed as Standard Effective Temperature (SET) (provide citations to Gagge/Foblets). SET is implemented in Std 55 as an index to quantify the equivalent temperature reduction caused by air movement cooling. The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated. This approach is intended to minimize errors introduced by the Fanger heat balance model’s reduced evaporative capacity on the predicted PMV result.
}
\todo[inline, color=blue!40]{
    FT. Firstly, it is important to note that the PMV does not aim to estimate sweating. 
    While this is a limitation, neither the PMV nor the PMV CE models include this capability.  
    The Std 55 did not modify the PMV heat balance equation but instead reduced the air and mean radiant temperatures by the cooling effect, assuming this adequately accounts for sweating. 
    However, this approach does not equate to an accurate heat balance calculation where sweat rate is explicitly modeled, as seen in the Gagge two-node model.  
    Consequently, the PMV CE does not: 1) use the Gagge two-node model to compute convective and evaporative heat loss from the body, nor 2) explicitly model physiological cooling in warm conditions where sweat evaporation becomes a key source of heat loss.  
    Secondly, in the PMV CE model, the cooling effect is applied in all conditions, not just in warm ones.  
    Thirdly, the definition of SET can also be somewhat unclear. 
    It is not calculated for a reference environment with standardised clothing and metabolic rates. 
    Instead, it represents the temperature of a hypothetical isothermal environment at 50\% RH, $<$0.1 m/s (20 fpm) average air velocity, where the total heat loss from the skin of an imaginary occupant (wearing clothing standardised for the activity) equals the heat loss from a person in the actual environment with their actual clothing and activity level.  
    Finally, as you mentioned, \textit{The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.} However, this does not guarantee that the person will not feel hot or sweat in this condition and, therefore, does not address the inherent issues of the original PMV model.
}

\subsection{\ac{pmv-ce} model}\label{subsec:pmv-ce-limitations}
The \gls{55} standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some changes to the \ac{pmv} inputs were implemented, nor for example, why the \ac{ce} is subtracted from the \ac{tdb} and \ac{tr}.
A partial justification of the model is mainly described in \mycite{arens_moving_2009} and secondarily in \mycite{yang_cooling_2015}.
However, we could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the \gls{55} over the original \ac{pmv} model.
\todo[inline]{EA. true enough, but in the standard, the objective is to have a rational physiology model.
The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
\todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are different, hence, publications about the latter do not sufficiently describe the rationale of the new model, moreover, the accuracy of the PMV CE was not tested}

@@

To illustrate the extent of the differences between the outputs of the two models, only as a function of \ac{vr}, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) assuming the same \ac{met}~=~\qty{1.2}{met} and \ac{clor}~=~\qty{0.5}{clo}.
We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
\todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. Many lab studies have measured results in that region.}
\todo[color=blue!40]{FT. In the manuscript, we decided not to include a Figure for 0.6m/s since in the ASHRAE DB 2 we hardly reach that value.
I am happy to generate a figure for you if need be}
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models for two values of \ac{vr}.
    We set \ac{met}~=~\qty{1.2}{met}, \ac{clor}~=~\qty{0.5}{clo}.
    \label{fig:comfort_regios_pmv_pmvce}}
\end{figure}
The results show that for \ac{vr}~=~\qty{0.4}{\m\per\s} and \ac{rh}~=~\qty{50}{\percent} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} and shifted towards warmer temperatures.
For \ac{vr}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.

@@

\begin{enumerate}
    \item Although the \ac{ce} claims to better account for the heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same \ac{pmv} model equations, hence it has the same limitations.
    This can be shown with a thought-example.
    Let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{vr}~=~\qty{0.25}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clor} of \qty{0.5}{clo}.
    In these conditions, a person will be sweating; in fact, according to Gagge's two-node model, which is used to calculate the \ac{set}, a person would have approximately \qty{30}{\percent} of their skin covered with sweat.
    The \ac{pmv} model, in this scenario, cannot accurately predict the \ac{tsv} since the person is sweating \mycite{huang_applicability_2014}.
    However, the only difference between the \ac{pmv} and \ac{pmv-ce} in this thought example is that the \ac{pmv-ce} accounts for the cooling effect due to the air movement (delta between \qty{0.25}{\m\per\s} and \qty{0.1}{\m\per\s}) which is equal to \qty{1.1}{\celsius}.
    Even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating.
    Therefore, in this scenario, where the \ac{pmv-ce} should excel compared to the \ac{pmv}, the \ac{pmv-ce} model still fails to accurately predict the \ac{tsv}.
    \todo[inline]{This is really confusing.
    You are evaluating the Cooling Effect in still air, when there is no Cooling Effect to evaluate.
    You are using as an example 29C in still air and 80\% RH which of course will cause sweating.
    You are citing skin wettedness values that have not been introduced in this paper (and are not part of any standard). 
    Where did they come from? You allude to results from Huang 2014 that the reader is supposed to know about. 
    The skin-wettedness discussion is irrelevant unless you go into a lot more physiological detail here.}
    \todo[inline, color=blue!40]{FT. We developed this thought-example to demonstrate that while the PMVCE seeks to improve the modeling of sweat rates, it does not alter the underlying PMV equations. 
    As a result, there are situations where the PMVCE may still fail to provide accurate predictions.
    We have modified the example and changed the air speed from 0.15 to 0.25 m/s}
    \item The previous example, highlights another limitation of the \ac{pmv-ce}.
    The assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is thermodynamically incorrect since the value of \ac{rh} is dependent on \ac{tdb}.
    The \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} value.
    \item It should be noted that over the years, the threshold value of \ac{vr} after which the \ac{ce} is calculated, has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
    However, these changes were not based on any scientific evidence, despite significantly affecting the output of the model.
    \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
    In some contexts, like in yearly building performance simulations, this could make a difference. 
    \todo[inline]{This may or may not be significant. If it is significant, the process could be simplified to a total of three calculations while remaining a rational method. What’s in there now was Tyler’s (CBE’s) choice. Charlie and I have used it more simply. We could propose this to the Standards Committee if we wanted to. We would have to calculate two validation tables for the two approaches.}
    \todo[inline, color=blue!40]{FT. I agree that this may or may not be significant, depending on the context and the computation speed could be improved.}
\end{enumerate}

>>>>

\subsection{\ac{pmv-ce} model}\label{subsec:pmv-ce-limitations}
The \gls{55} standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some changes to the \ac{pmv} inputs were implemented, nor for example, why the \ac{ce} is subtracted from the \ac{tdb} and \ac{tr}.
A partial justification of the model is mainly described in \mycite{arens_moving_2009} and secondarily in \mycite{yang_cooling_2015}.
However, we could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the \gls{55} over the original \ac{pmv} model.

@@

To illustrate the extent of the differences between the outputs of the two models, only as a function of \ac{vr}, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) assuming the same \ac{met}~=~\qty{1.2}{met} and \ac{clor}~=~\qty{0.5}{clo}.
We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models for two values of \ac{vr}.
    We set \ac{met}~=~\qty{1.2}{met}, \ac{clor}~=~\qty{0.5}{clo}.
    \label{fig:comfort_regios_pmv_pmvce}}
\end{figure}
The results show that for \ac{vr}~=~\qty{0.4}{\m\per\s} and \ac{rh}~=~\qty{50}{\percent} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} and shifted towards warmer temperatures.
For \ac{vr}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.

@@

\begin{enumerate}
    \item Although the \ac{ce} claims to better account for the heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same \ac{pmv} model equations, hence it has the same limitations.
    This can be shown with a thought-example.
    Let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{vr}~=~\qty{0.25}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clor} of \qty{0.5}{clo}.
    In these conditions, a person will be sweating; in fact, according to Gagge's two-node model, which is used to calculate the \ac{set}, a person would have approximately \qty{30}{\percent} of their skin covered with sweat.
    The \ac{pmv} model, in this scenario, cannot accurately predict the \ac{tsv} since the person is sweating \mycite{huang_applicability_2014}.
    However, the only difference between the \ac{pmv} and \ac{pmv-ce} in this thought example is that the \ac{pmv-ce} accounts for the cooling effect due to the air movement (delta between \qty{0.25}{\m\per\s} and \qty{0.1}{\m\per\s}) which is equal to \qty{1.1}{\celsius}.
    Even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating.
    Therefore, in this scenario, where the \ac{pmv-ce} should excel compared to the \ac{pmv}, the \ac{pmv-ce} model still fails to accurately predict the \ac{tsv}.
    \item The previous example, highlights another limitation of the \ac{pmv-ce}.
    The assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is thermodynamically incorrect since the value of \ac{rh} is dependent on \ac{tdb}.
    The \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} value.
    \item It should be noted that over the years, the threshold value of \ac{vr} after which the \ac{ce} is calculated, has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
    However, these changes were not based on any scientific evidence, despite significantly affecting the output of the model.
    \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
    In some contexts, like in yearly building performance simulations, this could make a difference. 
\end{enumerate}
